\subsection{Introduction}
\begin{frame}{Dimension Reduction Methods}{Introduction}

\begin{itemize}
    \item We now explore approaches that transform the predictors and then fit an OLS model using the transformed variables. \pause 

    \item Let $Z_1 , Z_2 , \cdots , Z_M$ represent $M < p$ linear combinations of our original reduction $p$ predictors. That is, \pause 

    \begin{equation}\label{eq:z-combs}
        Z_m = \sum_{j=1}^p \phi_{jm} X_j
    \end{equation} \pause 

    for some constants $\phi_{1m} , \phi_{2m} \cdots , \phi_{pm} , m = 1, \cdots , M$ . \pause We can then fit the model $y_i$ using OLS. \pause 

    \begin{equation}\label{eq:linear_dim}
        y_i = \theta_0 + \sum_{m=1}^M \theta_m z_{im} + \epsilon_i, \, \, \, i=1, \cdots, n. 
    \end{equation} \pause 
     
     \item The \textbf{dimension reduction} comes from reducing the problem of estimating the $p + 1$ coeﬃcients $\beta_p$ to estimating the $M + 1$ coeﬃcients $\theta_p$, where $M < p$.
\end{itemize}
    
\end{frame}

\begin{frame}{Dimension Reduction Methods}{Introduction}

Notice that from (\ref{eq:z-combs}), \pause 

\begin{equation*}
    \sum_{m=1}^M \theta_m z_{im} = \sum_{m=1}^M \theta_m \sum_{j=1}^p \phi_{jm} x_{ij} = \sum_{j=1}^p \sum_{m=1}^M \theta_m \phi_{jm} x_{ij} = \sum_{j=1}^p \beta_j x_{ij}, 
\end{equation*} \pause 
    where
    \begin{equation}\label{eq:beta-constrain}
        \beta_j = \sum_{m=1}^M \theta_m \phi_{jm}
    \end{equation} \pause 

\begin{itemize}
%    \item Dimension reduction serves to constrain the estimated $\beta_j$ coeﬃcients, since now they must take the form (\ref{eq:beta-constrain}) \pause 

    \item Where $p \gg n$, selecting $M \ll p$ can significantly reduce the variance of the fitted coeﬃcients. \pause 

    \item If $M = p$, and all the $Z_m$ are linearly independent, then (\ref{eq:beta-constrain}) poses no constraints. $\rightarrow$ OLS. \pause 

    \item The selection of the $\phi_{jm}$'s, can be achieved in diﬀerent ways: \pause 

    \begin{enumerate}
        \item Principal components. \pause 
        \item Partial least squares. 
    \end{enumerate}
\end{itemize}

    
\end{frame}

\subsection{Principal Component Regression}
\input{dim-reduction/pca.tex}

\subsection{Partial Least Squares}
\input{dim-reduction/pls.tex}